---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# systemfonts

<!-- badges: start -->
[![R build status](https://github.com/r-lib/systemfonts/workflows/R-CMD-check/badge.svg)](https://github.com/r-lib/systemfonts/actions)
[![Codecov test coverage](https://codecov.io/gh/r-lib/systemfonts/branch/master/graph/badge.svg)](https://codecov.io/gh/r-lib/systemfonts?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/systemfonts)](https://cran.r-project.org/package=systemfonts)
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
<!-- badges: end -->

systemfonts is a package that locates installed fonts. It uses the system-native libraries on Mac (CoreText) and Linux (FontConfig), and uses Freetype to parse the fonts in the registry on Windows.

## Installation
systemfonts is available from CRAN using `install.packages('systemfonts')`. It 
is however still under development and you can install the development version 
using devtools.

```{r, eval=FALSE}
# install.packages('devtools')
devtools::install_github('r-lib/systemfonts')
```

## Examples
The main use of this package is to locate font files based on family and style:

```{r}
library(systemfonts)

match_font('Avenir', italic = TRUE)
```

This function returns the path to the file holding the font, as well as the 
0-based index of the font in the file.

It is also possible to get a data.frame of all available fonts:

```{r, include=FALSE}
library(tibble) 
```
```{r}
system_fonts()
```

While getting this information in R is nice, the intended use is mostly through
compiled code so that graphic devices can easily locate relevant font files etc.

In order to use functions from systemfonts in C(++) code you should import them
using the `R_GetCCallable()` function, optimally wrapping it in a function to 
avoid repeated fetches, such as done by ragg:

```c++
static int locate_font(const char *family, int italic, int bold, char *path, int max_path_length) {
  static int (*p_locate_font)(const char *family, int italic, int bold, char *path, int max_path_length) = NULL;
  if (p_locate_font == NULL) {
    p_locate_font = (int(*)(const char *, int, int, char *, int)) R_GetCCallable("systemfonts", "locate_font");
  }
  return p_locate_font(family, italic, bold, path, max_path_length);
}
```

## System Defaults
systemfonts will always try to find a font for you, even if none exist with the
given family name or style. How it resolves this is system specific and should
not be relied on, but it can be expected that a valid font file is always 
returned no matter the input.

A few special aliases exist that behaves predictably but system dependent:

- `""` and `"sans"` return *Helvetica* on Mac, *Arial* on Windows, and the 
  default sans-serif font on Linux (*DejaVu Sans* on Ubuntu)
- `"serif"` return *Times* on Mac, *Times New Roman* on Windows, and the 
  default serif font on Linux (*DejaVu Serif* on Ubuntu)
- `"mono"` return *Courier* on Mac, *Courier New* on Windows, and the 
  default mono font on Linux (*DejaVu Mono* on Ubuntu)

## Code of Conduct
Please note that the 'systemfonts' project is released with a
[Contributor Code of Conduct](https://github.com/r-lib/systemfonts/blob/master/CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.
